\color{navyblue} Difference-in-Differences Estimation with Spatial Spillover

### 2021-03-22 latex/paper/Spillover.tex
Update introduction

    \begin{abstract}
        Empirical work often uses treatment variables defined by geographic boundaries. When researchers ignore the common problem that the effects of treatment cross over borders, classical difference-in-differences estimation produces biased estimates for the average treatment effect. In this paper, I decompose this bias in two parts. First, the control group no longer identifies the counterfactual trend because their outcomes are affected by treatment. Second, changes in treated units' outcomes reflect the effect of their own treatment status and the effect from the treatment status of ``close'' units. I propose estimation strategies that can remove both sources of bias as well as semi-parametrically estimate the spillovers themselves. Lastly, I extend this estimation strategy in an event-study framework following \citet{Callaway_Sant’Anna_2020}. Then, I turn to an empirical application revisiting an analysis of the Tennessee Valley Authority by \citet{Kline_Moretti_2014}. I highlight the importance of considering general equilibrium spillovers in the analysis of place-based policies when estimating the direct effect on targeted areas.
    \end{abstract}
\end{titlepage}
% ------------------------------------------------------------------------------

@@

Empirical work in economics often considers treatment assigned to groups of units, but the effect of these treatments potentially do not stay within these groups and spillover onto ‘close’ units.\footnote{The framework of this paper applies to any setting with a well-defined measures of distance, e.g. geographic
distance, economic distance such as supply chains, node distance in a graph, or social relationships in schools or cities.} For example, treatment is often assigned to geographic boundaries and spillovers occur from people crossing borders to access treatment (imperfect compliance) or from general equilibrium effects influencing neighboring areas. Despite the problem of spillovers being common across many settings, \citet{Berg_Streitz_2019} document that little empirical analysis directly controls for spillovers when estimating the `direct' effect of treatment on the treated units.\footnote{The term `direct' effect can refer to either the average treatment effect on the treated or the intent to treat effect. In empirical applications, the `direct` effect is sometimes called the `partial equilibrium' effect or the `local' effect.} The central result of my paper is that in the presence of spillovers, the difference-in-differences estimate identifies the direct effect of treatment (the estimand of interest) plus two additional bias terms resulting from the spillovers.

The intuition for the two bias terms resulting from spillovers is as follows. First, untreated units that are `close' to treated units experience effects of treatment and therefore these `control' units fail to identify the counterfactual trend. When estimating by difference-in-differences, the spillover onto the `close' control units is averaged into the untreated units' change in outcomes. In this case, the spillover is subtracted from the estimated treatment effect and biases the estimate in the opposite sign of the spillover effect. For example, consider trying to estimate employment effects of a factory opening in a county. Agglomeration economies suggest that neighboring counties would also benefit from knowledge spillovers and improved access to supply chain which would potentially increase employment \citep{Duranton_Puga_2003}. The treatment effect estimate is negatively biased because the change in outcome in neighboring counties is higher than it would be absent treatment. Researchers would therefore underestimate the positive benefits of treatment.\footnote{If the policy maker had a welfare function that included both treated and neighboring counties, they would fail to count the positive benefits of the control units and they would underestimate the positive benefits on the treated units. Difference-in-differences estimation would therefore result in double undercounting.}

Second, changes in treated units' outcomes reflect the effect of their own treatment status and the effect from the treatment status of ``close'' units. The spillover onto other treated units is averaged into the treated units' change in outcomes. Therefore the bias of treatment effect is the same sign as the sign of the spillover. Continuing with our example, two factory openings in neighboring counties might cause the benefit of each individual factory to increase due to agglomeration forces. The estimated treatment effect will count the `direct' effect of the factory opening as well as the benefit of the agglomeration economies. Therefore the treatment effect will be positively biased due to the positive spillover onto also treated units.

I then propose an estimation strategy that will remove both sources of bias with very minimal assumptions on the structure of spillovers. To remove \textit{all bias} from the direct effect estimate, a research needs to just include an indicator for being close to a treated unit interacted with treatment status, so long as the indicator captures all units affected by spillovers. An indicator removes all bias because in my decomposition of the treatment effect estimate, the bias terms are the \textit{average} spillover effects onto treated and control units. The interacted indicator will estimate this average and remove both bias terms from the treatment effect estimate. Importantly, my method does not require researchers to make any assumptions about how spillover effects propagate across space. The only assumption required is the maximum distance from treated units spillovers can occur.\footnote{Even if the maximum distance is not large enough, since treatment effects typically decay over distance, most of the bias will still be removed. However, including too many units will increase the variance in the estimates for average spillover effects.}

Since spillovers themselves are often important causal effects themselves, I next turn to how to estimate spillovers directly. I show with evidence from Monte Carlo simulations, that commonly used semi-parametric estimation strategies capture spillovers well (in a mean square prediction error sense). This involves creating a set of distance bins from treated units (e.g. being 0-20 miles, 20-40 miles, 40-60 miles from treated unit). The key consideration required by a researcher is whether the size of spillover effects are additive in the number of nearby treated units or not. For non-additive spillovers, I recommend a set of mutually-exclusive indicators for the closest treated unit being in each distance bin. For additive spillovers, I recommend using the number of treated units within each spillover bin.

As an example, I revisit the analysis of the Tennessee Valley Authority by \citet{Kline_Moretti_2014}. The Tennessee Valley Authority was a large scale New Deal program that created many new dams for flood-protection and navigation. The construction of large-scale dams lowered the cost of power for industrial firms giving the region an industrial boost \citep{Kitchens_2014}. The scale of federal investment in the region was large and the pro-manufacturing benefits likely spread further than the Authority's boundary due to the electrification infrastucture and agglomeration economies \citet{Severnini_2014}. I show that estimation by difference-in-differences fails to account for these spillovers and for different outcome variable it underestimates and overestimates the local effect, depending on the sign of the spillovers.

I turn to an empirical application in the literature in urban economics on local place-based policies in section \ref{sec:tva}. Using structural urban models, \citet{Kline_Moretti_2014b} show that the `local' or partial equilibrium effects of a place-based policy are offset in part by indirect or general equilibrium effects in non-treated locations. I contribute by showing that because of these general equilibrium effects, estimation of local effects are often underestimated by not controlling for the spillover effects in estimation.

Last, in section \ref{sec:event_study}, I extend estimation of the direct effect and spillover effects of treatment into the event study framework by extending the work of \citet{Callaway_Sant’Anna_2020}. I first show how to adjust their methods to control for spatial spillovers in estimation of treatment effects. Second, I show how an extension of their method allows for estimation of spillover effects. My paper is the first paper to study estimation of treatment effects in an event-study framework.

@@

More recently, a new strand of literature exploring estimation of direct and spillover effects which does not require a completely untreated cluster by using a potential outcomes framework. \citet{Vazquez-Bare_2019} presents a potential outcomes model that explicitly accounts for ``within-group'' spillovers in experiments which allows for seperate estimation of direct effect and spillover effects by a simple differences-in-means. I extend this work by focusing on difference-in-differences estimation in non-experimental settings.  \citet{Sävje_Aronow_Hudgens_2019} consider estimation of treatment effects by difference-in-differences, but they define treatment effects as the combination of direct effects and spillover effects. My paper develops a strategy to allow researchers to seperately identify treatment effects and spillovers.These estimates can be combined afterwards to estimate their definition of average treatment effect. The further advantage is that `net' treatment effects can be estimated at different levels of exposure which is more relevant for future policy decisions.

@@

The rest of the paper is structured as follows. Section \ref{sec:po_framework} presents the potential outcomes framework, defines the estimand of interest, and shows the resulting bias from estimating a classical difference-in-differences model. Section \ref{sec:estimation} presents Monte Carlo simulations to illlustrate the bias result and to evaluate currently used solutions in the literature. Section \ref{sec:tva} presents an application for evaluating the Tennessee Valley Authority program. Section \ref{sec:event_study} discusses briefly how to incorporate spillovers into event-study estimation following the insights from \citet{Callaway_Sant’Anna_2020}.
% @TODO Fix this

>>>>

    \begin{abstract}
        Empirical work often uses treatment variables defined by geographic boundaries. When researchers ignore the common problem that the effects of treatment cross over borders, classical difference-in-differences estimation produces biased estimates for the average treatment effect. In this paper, I decompose this bias in two parts. First, the control group no longer identifies the counterfactual trend because their outcomes are affected by treatment. Second, changes in treated units' outcomes reflect the effect of their own treatment status and the effect from the treatment status of ``close'' units. I propose estimation strategies that can remove both sources of bias as well as semi-parametrically estimate the spillovers themselves. Lastly, I extend this estimation strategy in an event-study framework following \citet{Callaway_Sant’Anna_2020} to allow for staggered treatment adoption. Then, I turn to an empirical application revisiting an analysis of the Tennessee Valley Authority by \citet{Kline_Moretti_2014}. I highlight the importance of considering general equilibrium spillovers in the analysis of place-based policies when estimating the direct effect on targeted areas.
    \end{abstract}
\end{titlepage}
% ------------------------------------------------------------------------------

@@

Empirical work in economics often considers settings where treatment is assigned to groups of units, but the effect of these treatments spillover onto ‘close’ units.\footnote{The framework of this paper applies to any setting with a well-defined measures of distance, e.g. geographic
distance, economic distance such as supply chains, node distance in a graph, or social relationships in schools or cities.} For example, treatment is commonly assigned to geographic boundaries and spillovers occur from people crossing borders to access treatment (i.e. imperfect compliance) or from general equilibrium effects influencing neighboring areas. Despite the problem of spillovers being common across many settings, \citet{Berg_Streitz_2019} document that little empirical analysis directly controls for spillovers when estimating the `direct' effect of treatment on the treated units. The term `direct' effect can refer to either the average treatment effect on the treated or the intent to treat effect.\footnote{In empirical applications, the `direct` effect is sometimes called the `partial equilibrium' effect or the `local' effect.} The central theoretical result of my paper is that in the presence of spillovers, the standard difference-in-differences estimate identifies the direct effect of treatment (the estimand of interest) plus two additional bias terms resulting from the spillovers.

The intuition for the two bias terms resulting from spillovers is as follows. First, untreated units that are `close' to treated units experience effects of treatment and therefore these `control' units fail to identify the counterfactual trend. When estimating by difference-in-differences, the spillover onto the `close' control units is averaged into the untreated units' change in outcomes. In this case, the spillover is subtracted from the estimated treatment effect and biases the estimate in the opposite sign of the spillover effect. For example, consider trying to estimate employment effects of a factory opening in a county. Agglomeration economies suggest that neighboring counties would also benefit from knowledge spillovers and improved access to supply chain which would potentially increase economic activity \citep{Duranton_Puga_2003}. The treatment effect estimate is negatively biased because the change in outcome in neighboring counties is higher than it would be absent treatment. Researchers would therefore underestimate the positive benefits of treatment.\footnote{If the policy maker had a welfare function that included both treated and neighboring counties, they would fail to count the positive benefits of the control units and they would underestimate the positive benefits on the treated units. Difference-in-differences estimation would therefore result in double undercounting.}

% @TODO: Improve examples to better build intuition

Second, changes in treated units' outcomes reflect the effect of their own treatment status and the effect from the treatment status of ``close'' units. The spillover onto other treated units is averaged into the treated units' change in outcomes. Therefore the bias of treatment effect is the same sign as the sign of the spillover. Continuing with our example, two factory openings in neighboring counties might cause the benefit of each individual factory to increase due to agglomeration forces. The estimated treatment effect will count the `direct' effect of the factory opening as well as the benefit of the agglomeration economies from nearby factory openings. Therefore the treatment effect will be positively biased due to the positive spillover onto also treated units.

I then use my framework to evaluate commonly used empirical strategies found in applied work and provide practical recommendations for researchers. The most common approach is to remove potentially contaminated control units from the sample in order to remove spillover effects on control units \citep{Berg_Streitz_2019}. This approach is not recommended for two reasons. First, the removal of too few control units will leave bias and removal of too many control units will lower precision of the treatment effect estimate. This results in a bias-variance trade-off that other approaches do not share. Second, this method does not remove spillover effects on treated units and will leave this bias term in the difference-in-differences estimate. Another common approach is to include either an indicator or a set of rings around treated units. This approach will remove all bias from the spillover effects on control units provided that all control units with spillover effects are included in the indicator. Similar to the previous approach, spillover effects on also treated units still add bias to estimates of the direct effect.

Therefore, I propose an estimation strategy that will remove both sources of bias with very minimal assumptions on the structure of spillovers. To remove \textit{all bias} from the direct effect estimate, a research needs to just include an indicator for being close to a treated unit interacted with treatment status, so long as the indicator captures all units affected by spillovers. In my decomposition of the treatment effect estimate, the bias terms are the \textit{average} spillover effects onto treated and control units. An idicator variable interacted with treatment status therefore can estimate these two averages and remove them from the difference-in-differences estimate. Importantly, my method does not require researchers to make any assumptions about how spillover effects propagate across space. The only assumption required is the maximum distance from treated units spillovers can occur.\footnote{Even if the maximum distance is not large enough, since treatment effects typically decay over distance, most of the bias will still be removed. However, including too many units will increase the variance in the estimates for average spillover effects.}

Since spillover effects are often important causal effects themselves, I next turn to how to estimate spillovers directly. I show with evidence from Monte Carlo simulations, that commonly used semi-parametric estimation strategies capture spillovers well (in a mean square prediction error sense). This involves creating a set of distance bins from treated units (e.g. being 0-20 miles, 20-40 miles, 40-60 miles from treated unit) and interacting them with a treatment indicator.\footnote{The choice of the distance bins depends on the economic context and in particular the source of the spillovers.
% TODO: Reference Butts and Clarke
} The key consideration required by a researcher is whether the size of spillover effects are additive in the number of nearby treated units or not. For non-additive spillovers, I recommend a set of mutually-exclusive indicators which measures which distance bin that the closest treated unit falls within for a given observation. For additive spillovers, I recommend using the number of treated units within each distance bin for a given observation.\footnote{While the non-additive version of ring indicators will remove all bias from the direct effect of treatment even under misspecification, this is not true of additive rings. Therefore, researchers may want to estimate the direct effect of treatment using non-additive rings and then use additive rings to estimate spillover effects.}

I turn to an empirical application in the literature in urban economics on local place-based policies in Section \ref{sec:tva}. Using structural urban models, \citet{Kline_Moretti_2014b} show that the `local' or partial equilibrium effects of a place-based policy are offset in part by indirect or general equilibrium effects in non-treated locations. In particular, their model implies these effects are declining in distance. I contribute by showing that because of these general equilibrium effects, estimation of local effects are often misestimated by not controlling for the spillover effects in their empirical specification.

As an example, I revisit the analysis of the Tennessee Valley Authority by \citet{Kline_Moretti_2014}. The Tennessee Valley Authority was a large scale New Deal program that created many new dams for flood-protection and navigation. The construction of large-scale dams lowered the cost of power for industrial firms giving the region an industrial boost \citep{Kitchens_2014}. The scale of federal investment in the region was large and the pro-manufacturing benefits likely spread further than the Authority's boundary due to the electrification infrastucture and agglomeration economies \citep{Severnini_2014}. I show that estimation by difference-in-differences fails to account for these spillovers and for different outcome variable it underestimates and overestimates the local effect, depending on the sign of the spillovers.

Following this empirical application, I discuss how my framework fits into a larger discussion on identification strategies with place-based policies. I revisit conflicting results about United States' federal Empowerment Zones, a program that creates incentives for businesses to locate in high-poverty neighborhoods. \citet{Busso_Gregory_Kline_2013} use Census Tracts from qualified but ultimately rejected applications that are typically far away from accepted Zones as a comparison group and find that significant reductions in poverty. This empirical strategy is based on the idea that since they qualified for the program they would be on similar trends without being contaminated by spillover effects. \citet{Neumark_Kolko_2010}, on the other hand, use census tracts within 1,000 feet of the Zone and find no statistically significant effects. This empirical strategy assumes that the borders are drawn somewhat randomly and therefore being just outside is as good as random. Although, using close units is potentially problematic as they can experience spillover effects from the Zones. My framework can explain the differences in findings if there are positive spillover effects onto census tracts just outside Empowerment Zones.

Last, in Section \ref{sec:event_study}, I extend estimation of the direct effect and spillover effects of treatment into the event study framework by extending the work of \citet{Callaway_Sant’Anna_2020}. This allows for a very common setting in which treatment turns on for different units in different periods. I first show how to adjust their methods to control for spatial spillovers in estimation of treatment effects. Then, I show how to estimate average spillover effects. My paper is the first paper to study estimation of treatment effects in an event-study framework in the presence of spillover effects.

@@

There is a nascent literature exploring estimation of direct and spillover effects which does not require a completely untreated cluster by using a potential outcomes framework. \citet{Vazquez-Bare_2019} presents a potential outcomes model that explicitly accounts for ``within-group'' spillovers in experiments which allows for seperate estimation of direct effect and spillover effects by a simple differences-in-means. I extend this work by focusing on difference-in-differences estimation in non-experimental settings.  \citet{Sävje_Aronow_Hudgens_2019} consider estimation of treatment effects by difference-in-differences, but they define treatment effects as the combination of direct effects and spillover effects. My paper develops a strategy to allow researchers to seperately identify treatment effects and spillovers. These estimates can be combined to estimate their definition of average treatment effect. The further advantage is that `net' treatment effects can be estimated at different levels of exposure which is more relevant for future policy decisions.

@@

The rest of the paper is structured as follows. Section \ref{sec:po_framework} presents the potential outcomes framework, defines the estimand of interest, and shows the resulting bias from estimating a classical difference-in-differences model. Section \ref{sec:estimation} presents Monte Carlo simulations to illlustrate the bias result and to evaluate currently used solutions in the literature. Section \ref{sec:tva} presents an application for evaluating the Tennessee Valley Authority program. Section \ref{sec:event_study} discusses briefly how to incorporate spillovers into event-study estimation following the insights from \citet{Callaway_Sant’Anna_2020}.
